---
title: "Positive Covid-19 Test Rate Increases Around DC"
output:
  rmarkdown::github_document: default
  html_document: 
    keep_md: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(lubridate)
library(readxl)
library(httr)

setwd("..")

sheet_url <- "https://coronavirus.dc.gov/sites/default/files/dc/sites/coronavirus/page_content/attachments/DC-COVID-19-Data-for-April-26-2020.xlsx"

GET(sheet_url, write_disk(tf <- tempfile(fileext = ".xlsx")))

overall_dc <- read_excel(tf, sheet = "Overal Stats")
colnames(overall_dc)[1:2] <- c("Category", "Subcategory")
overall_dc <- overall_dc %>% dplyr::filter(!is.na(Category))

### Testing data missing from 20 March, so averaging 19 March and 21 March data
testing <- overall_dc %>% dplyr::filter(Category == "Testing")
for(i in 3:ncol(testing)) testing[i] <- testing[[i]] %>% replace_na(0)
for(i in 3:ncol(testing)) testing[i] <- as.numeric(testing[[i]])

testing_test <- testing %>% dplyr::select(-Category) %>% 
  pivot_longer(-Subcategory, names_to = "Date", values_to = "Count")
testing_test$Date <- as.numeric(testing_test$Date)
testing_test$Date <-  as.Date(testing_test$Date - 2, origin = "1900-01-01")

testing_test <- testing_test %>% pivot_wider(names_from = Subcategory,
                                             values_from = Count)
colnames(testing_test)[2:5] <- c("Tested_Overall", "Tested_Positive",
                                 "Deaths_Cuml", "Recovered_Cuml")

### Fix the missing value for Tested_Overall on 20 March
testing_test[testing_test$Date == "2020-03-20", 2] <- 814

testing_test <- testing_test %>% 
  mutate(Test_New = Tested_Overall - lag(Tested_Overall, default = 0),
         Pos_New = Tested_Positive - lag(Tested_Positive, default = 0),
         Test_Lag_1 = lag(Test_New, n = 1, default = 0),
         Test_Lag_2 = lag(Test_New, n = 2, default = 0),
         Test_Lag_3 = lag(Test_New, n = 3, default = 0),
         Test_Lag_4 = lag(Test_New, n = 4, default = 0),
         Test_Lag_5 = lag(Test_New, n = 5, default = 0),
         Test_Lag_6 = lag(Test_New, n = 6, default = 0),
         Test_Lag_7 = lag(Test_New, n = 7, default = 0),
         Test_Lag_8 = lag(Test_New, n = 8, default = 0),
         Test_Lag_9 = lag(Test_New, n = 9, default = 0),
         Test_Lag_10 = lag(Test_New, n = 10, default = 0)
         )

covid_tracking <- read_csv("https://covidtracking.com/api/v1/states/daily.csv") %>%
  dplyr::filter(state %in% c("DC", "MD", "VA", "NY",
                             "NJ", "MI", "LA")) %>% 
  dplyr::rename(State = state) %>% 
  group_by(State) %>% 
  dplyr::mutate(date = as_date(as.character(date),
                               format = "%Y%m%d", tz = "UTC")) %>% 
  dplyr::mutate(pos_rate = positive / (positive + negative) * 100,
                pos_tests_3day = positiveIncrease + 
                  lag(positiveIncrease, n = 1L, order_by = date) +
                  lag(positiveIncrease, n = 2L, order_by = date),
                total_tests_3day = totalTestResultsIncrease + 
                  lag(totalTestResultsIncrease, n = 1L, order_by = date)+
                  lag(totalTestResultsIncrease, n = 2L, order_by = date),
                pos_rate_3day = pos_tests_3day / 
                  total_tests_3day * 100,
                death_in_3 = lead(deathIncrease, n = 3L,
                                  order_by = date),
                # Population estimates from Census 2019
                population = case_when(
                  State == "DC" ~ 705749,
                  State == "LA" ~ 4648790,
                  State == "MI" ~ 9986860,
                  State == "MD" ~ 6045680,
                  State == "VA" ~ 8535519,
                  State == "NY" ~ 19453561,
                  State == "NJ" ~ 8882190
                ),
                deathIncPerMil = deathIncrease/population * 1000000) %>% 
  dplyr::filter(!(State == "MD" & date < "2020-03-28"))

current_rates <- dplyr::filter(covid_tracking, date == max(date)) %>% 
  select(pos_rate, pos_rate_3day)
```

`r Sys.time()`

The number of Covid-19 tests conducted in the DC area increased 
considerably through the first part of April and has generally 
plateaued since then. However, the positive test rate has increased
recently, suggesting that the plateauing case numbers may be a result
of plateauing testing capability.

The graph below shows a rolling three-day average of tests conducted in
DC, Maryland, and Virginia. Note: Maryland only began reporting negative
tests in late March, so I've truncated most of the Maryland data until it
doesn't show abnormal spikes in late March.

```{r dmv_testing_numbers}
ggplot(data = dplyr::filter(covid_tracking,
                            State %in% c("DC", "MD", "VA"),
                            !(State == "MD" & date < "2020-03-31")),
       aes(x = date, y = total_tests_3day / 3, color = State)) +
  geom_line(size = 1) + 
  labs(x = "Date", y = "Covid-19 Tests Conducted\n3 Day Average",
       title = "Covid-19 Tests Completed Near DC",
       caption = "Data from Covid Tracking Project")
```

Despite plateauing or decreasing numbers of tests in the area, the 
cumulative positive test rate has increased over time since the rates 
spiked in mid-March, when fewer Covid-19 tests were being conducted.
Right now, the cumulate positive rate in DC is 
`r round(current_rates[current_rates$State == "DC", "pos_rate"][[1]], 1)`%, the rate in Maryland is 
`r round(current_rates[current_rates$State == "MD", "pos_rate"][[1]], 1)`%, and the rate in Virginia is
`r round(current_rates[current_rates$State == "VA", "pos_rate"][[1]], 1)`%.  

```{r pos_rate_cumulative}
ggplot(data = dplyr::filter(covid_tracking,
                            State %in% c("DC", "MD", "VA")),
       aes(x = date, y = pos_rate, color = State)) +
  geom_line(size = 1) + 
  labs(x = "Date", y = "Cumulative Positive Test Rate (%)",
       title = "Covid-19 Positive Test (Cumulative) Rate Near DC",
       caption = "Data from Covid Tracking Project")
```

The recent increases in positive test rates are more pronounced 
when displayed as recent, rather than cumulative, averages. The average positive test rate between `r Sys.Date()-3` and `r Sys.Date()-1` was 
`r round(current_rates[current_rates$State == "DC", "pos_rate_3day"][[1]], 1)`% in DC,
`r round(current_rates[current_rates$State == "MD", "pos_rate_3day"][[1]], 1)`% in Maryland, and `r round(current_rates[current_rates$State == "VA", "pos_rate_3day"][[1]], 1)`% in Virginia.  
`
```{r rolling_pos_rate}
ggplot(data = dplyr::filter(covid_tracking,
                            State %in% c("DC", "MD", "VA")),
       aes(x = date, y = pos_rate_3day, color = State)) +
  geom_line(size = 1) + 
  labs(x = "Date", y = "Rolling 3 Day Positive Test Rate (%)",
       title = "Covid-19 Positive Test Rate Near DC",
       caption = "Data from Covid Tracking Project") +
  ylim(0, 40) +
  xlim(min(covid_tracking$date) + 10, max(covid_tracking$date))
```

New York state has generally had the highest testing rate of any state,
due to New York City being the center of the Covid-19 pandemic in the 
United States. However, even with a high number of tests, the positive 
test rate spiked precipitously as the daily death rate started climbing 
in the state. The positive test rate peaked more than a week before the 
daily death rate peaked in the state.

```{r ny_plot}
ggplot(data = dplyr::filter(covid_tracking,
                            State %in% c("NY")), aes(x = date)) + 
  geom_line(aes(y = deathIncPerMil), size = 1, color = "red") +
  geom_line(aes(y = pos_rate_3day), size = 1, color = "blue") +
  xlim(min(covid_tracking$date) + 10, max(covid_tracking$date)) +
  ylim(0, 60) + 
  labs(title = "Covid-19 Rates in NY State",
       x = "Date", y = "Rate",
       caption = "Data from Covid Tracking Project") +
  annotate("text", label = "Positive Test Rate - 3 Day Average",
           x = max(covid_tracking$date) - 23,
           y = 57, color = "blue") +
  annotate("text", label = "New Deaths per million", color = "red",
           x = max(covid_tracking$date) - 18, y = 10)
```

The DC area clearly hasn't seen a rise like in New York. However, the 
rising positive test rate and plateauing number of tests that are being
conducted in DC, Maryland, and Virginia suggests that the area probably
isn't experiencing a major decline in the number of cases. Also, after
that statement, it's probably a good time to mention I am not an 
epidemiologist and probably talk to one or read something by one before 
taking that previous sentence as gospel.

## Appendix: DC Test Rates

This plot simply shows the cumulative positive test rate using data 
directly from the DC government.

```{r dc_data}
plot3 <- testing_test %>% 
  dplyr::filter(Date >= "2020-03-13") %>% 
  mutate(pos_rate_cumulative = Tested_Positive / 
           Tested_Overall * 100)

gg_rate <- ggplot(data = plot3, aes(x = Date, y = pos_rate_cumulative )) +
  geom_line() + ylab("Cumulative Positive Test Rate") + 
  ylim(c(0, max(plot3$pos_rate_cumulative, na.rm = T) + 2)) + 
  ggtitle("DC Positive Covid-19 Test Rate Over Time") +
  theme(plot.title = element_text(hjust = 0.5))
gg_rate
```