---
title: "Lag in DC Test Results"
output:
  rmarkdown::github_document: default
  html_document: 
    keep_md: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(readxl)
library(dplyr)
library(tidyr)
library(ggplot2)
library(httr)

setwd("..")

sheet_url <- "https://coronavirus.dc.gov/sites/default/files/dc/sites/coronavirus/page_content/attachments/DC-COVID-19-Data-for-May-4-2020.xlsx"

GET(sheet_url, write_disk(tf <- tempfile(fileext = ".xlsx")))

overall_dc <- read_excel(tf, sheet = "Overal Stats")
colnames(overall_dc)[1:2] <- c("Category", "Subcategory")
overall_dc <- overall_dc %>% dplyr::filter(!is.na(Category))

### Testing data missing from 20 March, so averaging 19 March and 21 March data
testing <- overall_dc %>% dplyr::filter(Category == "Testing")
for(i in 3:ncol(testing)) testing[i] <- testing[[i]] %>% replace_na(0)
for(i in 3:ncol(testing)) testing[i] <- as.numeric(testing[[i]])

testing_test <- testing %>% dplyr::select(-Category) %>% 
  pivot_longer(-Subcategory, names_to = "Date", values_to = "Count")
testing_test$Date <- as.numeric(testing_test$Date)
testing_test$Date <-  as.Date(testing_test$Date - 2, origin = "1900-01-01")

testing_test <- testing_test %>% pivot_wider(names_from = Subcategory,
                                             values_from = Count)
colnames(testing_test)[2:5] <- c("Tested_Overall", "Tested_Positive",
                                 "Deaths_Cuml", "Recovered_Cuml")

### Fix the missing value for Tested_Overall on 20 March
testing_test[testing_test$Date == "2020-03-20", 2] <- 814

testing_test <- testing_test %>% 
  mutate(Test_New = Tested_Overall - lag(Tested_Overall, default = 0),
         Pos_New = Tested_Positive - lag(Tested_Positive, default = 0),
         Test_Lag_1 = lag(Test_New, n = 1, default = 0),
         Test_Lag_2 = lag(Test_New, n = 2, default = 0),
         Test_Lag_3 = lag(Test_New, n = 3, default = 0),
         Test_Lag_4 = lag(Test_New, n = 4, default = 0),
         Test_Lag_5 = lag(Test_New, n = 5, default = 0),
         Test_Lag_6 = lag(Test_New, n = 6, default = 0),
         Test_Lag_7 = lag(Test_New, n = 7, default = 0),
         Test_Lag_8 = lag(Test_New, n = 8, default = 0),
         Test_Lag_9 = lag(Test_New, n = 9, default = 0),
         Test_Lag_10 = lag(Test_New, n = 10, default = 0)
         )
```

### Postive Covid-19 Results in DC Lag by About Four Days

`r format(Sys.time(), '%d %B, %Y')`

**Note: I'm going to stop updating this file as the clear relationship
between tests a few days ago and positive numbers now doesn't appear to
be valid anymore.**

The below plot shows the number of Positive Tests in red and the number of New
Tests (multiplied by the positive test rate) in blue. Although the numbers 
tracked pretty well early on in testing, they have diverged recently. The 
correlation coefficient is currently `r cor(testing_test[6:ncol(testing_test)])[2, 1]`.

```{r same_day}
plot1 <- testing_test %>% 
  dplyr::mutate(Test_by_Pos_Rate = Test_New / (max(Tested_Overall)/max(Tested_Positive))) %>% 
  dplyr::select(Date, Pos_New, Test_by_Pos_Rate) %>% 
  tidyr::gather(key = "Test_Metric", value = "Number",
                Pos_New, Test_by_Pos_Rate)

ggplot(data = plot1, aes(x = Date)) +
  geom_line(aes(y = Number, color = Test_Metric), size = 1) +
  labs(title = "Positive Tests and Total Tests for Covid-19 in DC", y = "Tests", x = "Date") +
  scale_color_discrete(name = "Testing Metric by Day",
                       labels = c("Positve Tests",
                                  "Total Tests\n(times cumulative\npositive test rate)")) +
  theme(plot.title = element_text(hjust = 0.5))
```

The below plot shows the number of Positive Tests in red compared to
the number of New Tests four days ago (times the cumulative positive test rate) 
in blue. These numbers track much better, with a current correlation coefficient
of `r cor(testing_test[6:ncol(testing_test)])[2, 6]`.

```{r lag_4}
plot2 <- testing_test %>% 
  dplyr::mutate(Test_Lag_4_by_Pos_Rate = Test_Lag_4 / (max(Tested_Overall)/max(Tested_Positive))) %>% 
  dplyr::select(Date, Pos_New, Test_Lag_4_by_Pos_Rate) %>% 
  tidyr::gather(key = "Test_Metric", value = "Number",
                Pos_New, Test_Lag_4_by_Pos_Rate)

ggplot(data = plot2, aes(x = Date)) +
  geom_line(aes(y = Number, color = Test_Metric), size = 1) +
  labs(title = "Positive Tests and Lagged Total Tests for Covid-19 in DC",
       y = "Tests", x = "Date") +
  scale_color_discrete(name = "Testing Metric by Day",
                       labels = c("Positve Tests",
                                  "Total Tests 4 Days Ago\n(times cumulative\n positive test rate)")) +
  theme(plot.title = element_text(hjust = 0.5))
```

Here is the correlation matrix for those interested in such a thing.

```{r}
cor(testing_test[6:ncol(testing_test)])
```

A couple notes. First, the test counts for 20 March were missing so I averaged 
the 19 March and 21 March test rate. In addition, for the initial lagged 
variables, I made the defaults zero. This could make the correlation metrics
stronger for the lagged variables, but the effect doesn't look uniform.

### Cumulitave Positive Test Rate

```{r}
plot3 <- testing_test %>% 
  dplyr::filter(Date >= "2020-03-13") %>% 
  mutate(pos_rate_cumulative = Tested_Positive / 
           Tested_Overall * 100)

gg_rate <- ggplot(data = plot3, aes(x = Date, y = pos_rate_cumulative )) +
  geom_line() + ylab("Cumulative Positive Test Rate") + 
  ylim(c(0, max(plot3$pos_rate_cumulative, na.rm = T) + 2)) + 
  ggtitle("DC Positive Covid-19 Test Rate Over Time") +
  theme(plot.title = element_text(hjust = 0.5))
gg_rate
```

